Value Ordering, Discrepancies, and Restarts in Subgraph Solvers: \\ Some Half-Baked Mucking Around

### 2017-10-01 paper/paper.tex
Tweaks

\documentclass{article}
\usepackage{ijcai17}
\usepackage{times}
\usepackage{microtype}
\usepackage{amsmath}
\usepackage[ruled,vlined]{algorithm2e}
\usepackage{tikz}
\usepackage{cleveref}

@@

@@

@@

The performance of the biased and tailored heuristics is very similar: the biased heuristic has
slightly worse performance below a 100s timeout, and effectively equal performance above it. This
shows that we can introduce an element of randomness into the tailored value-ordering heuristic
without adversely affecting its performance.

    \caption{Cumulative number of instances solved over time, using the four different
    value-ordering heuristics.}
    \label{figure:value-ordering-heuristics}
\end{figure}

@@

In \cref{figure:restarts} we show the effects of adding restarts to the algorithm. With restarts,
the random value-ordering heuristic comfortably beats the tailored strategy without restarts. In
other words, although having a good variable-ordering heuristic is beneficial, introducing
randomness into the search is better, if it is done alongside a mechanism to avoid heavy commitment
to any particular random choice.  However, the biased heuristic together with restarts is better
still---that is, if we are introducing restarts, then it is better to add a small amount of
randomness to a tailored deterministic heuristic than it is to throw away the heuristic altogether.

@@

@@

    \bigskip
    \centering
    \includegraphics*{gen-graph-scatter-random.pdf}

    \caption{Instance by instance comparison of the basic algorithm with the tailored heuristic,
    versus biased + restarts. Below, comparing instead with simple random + restarts.}
    \label{figure:scatter}
\end{figure}

Cumulative plots only show aggregate behaviour. What happens on an instance by instance basis?  In
the top of \cref{figure:scatter} we scatter plot performance, showing satisfiable and unsatisfiable
instances using dots and crosses respectively. On the $x$-axis we show the tailored heuristic,
whilst on the $y$-axis we show biased + restarts. Points on the outer axes represent instances
which timed out with one heuristic but not the other.

Nearly all of the unsatisfiable instances are on or slightly above the $x-y$ diagonal, showing that
their performance is either unchanged or reduced slightly. On the other hand, there are ?? large
numbers of satisfiable instances well below the diagonal line, indicating large speedups. Better
yet, there are only ?? satisfiable instances that are more than a factor of ?? times worse.

@@

For contrast, the bottom plot of \cref{figure:scatter} compares the tailored heuristic with random +
restarts. The performance on unsatisfiable instances is the same. However, although better in
aggregate, we see many more satisfiable instances performing substantially worse. ?? Quantify this a
bit.

>>>>

\documentclass{article}
\usepackage{ijcai17}
\usepackage{times}
\usepackage{microtype}
\usepackage{amsmath}
\usepackage[ruled,vlined]{algorithm2e}
\usepackage{tikz}

\usepackage{hyperref}
\usepackage{cleveref}

@@

    \bigskip

@@

    \caption{The cumulative number of instances solved over time, using the four different
    value-ordering heuristics.}
    \label{figure:value-ordering-heuristics}
\end{figure}

@@

\begin{figure}[tb]
    \centering
    \includegraphics*{gen-graph-scatter-heuristics.pdf}

    \caption{An instance by instance comparison of the tailored and biased heuristics.}
    \label{figure:scatter-heuristics}
\end{figure}

The performance of the biased and tailored heuristics is very similar: the biased heuristic has
slightly worse performance below a 100s timeout, and effectively equal performance above it. This
shows that we can introduce an element of randomness into the tailored value-ordering heuristic
without adversely affecting its performance \emph{in aggregate}. \Cref{figure:scatter-heuristics}
gives a more detailed look at these results. In this plot, satisfiable and unsatisfiable instances
are shown using dots and crosses respectively. On the $x$-axis we measure the tailored heuristic,
whilst on the $y$-axis we measure random + restarts. Points on the outer axes represent instances
which timed out with one heuristic but not the other. This plot shows that on a case by case basis,
randomness can make a large difference to individual satisfiable instances.

@@

In \cref{figure:restarts} we show the effects of adding restarts to the algorithm. With restarts,
the random value-ordering heuristic comfortably beats the tailored strategy without restarts. In
other words, although having a good variable-ordering heuristic is beneficial, introducing
randomness into the search is better, if it is done alongside a mechanism to avoid heavy commitment
to any particular random choice. \Cref{figure:scatter-random} looks at this in more detail.  The
performance on unsatisfiable instances is the same. However, although better in aggregate, we see
many satisfiable instances performing substantially worse. ?? Quantify this a bit.

@@

\begin{figure}[p]
    \centering
    \includegraphics*{gen-graph-scatter-random.pdf}
    \caption{Instance by instance comparison of the basic algorithm with the tailored heuristic,
    versus random + restarts.}
    \label{figure:scatter-random}
\end{figure}

@@

    \caption{Instance by instance comparison of the basic algorithm with the tailored heuristic,
    versus biased + restarts.}
    \label{figure:scatter}
\end{figure}

\Cref{figure:restarts} shows that the biased heuristic together with restarts is better
still---that is, if we are introducing restarts, then it is better to add a small amount of
randomness to a tailored deterministic heuristic than it is to throw away the heuristic altogether.
In the more detailed view in \cref{figure:scatter},
nearly all of the unsatisfiable instances are on or slightly above the $x-y$ diagonal, showing that
their performance is either unchanged or reduced slightly. On the other hand, there are ?? large
numbers of satisfiable instances well below the diagonal line, indicating large speedups. Better
yet, there are only ?? satisfiable instances that are more than a factor of ?? times worse.

?? This has made up the consistency we lost from introducing randomness.

@@
